---
title: "Working with models"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Working with models}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```
{lvmisc} has a set of tools to work with model objects. Currently, the supported model classes are `lm` from the {stats} package and `lmerMod` from the {lme4} [package](https://CRAN.R-project.org/package=lme4). If you would like to have methods implemented to other model classes, please file an issue at https://github.com/verasls/lvmisc/issues.

The functions to work with models can be divided into three categories:

1. Model diagnostics
1. Model cross-validation
1. Model accuracy

Before starting, let's load the required packages:

```{r setup, message = FALSE}
library(lvmisc)
library(lme4)
```

# Model diagnostics

{lvmisc} has a plotting method for some common model diagnostic plots. To explore it, let's first create some example models:

```{r diagnostic_models}
m1 <- lm(disp ~ mpg + hp + cyl + mpg:cyl, mtcars)
m2 <- lmer(disp ~ mpg + (1 | cyl), mtcars)
```
